Refinements for open automata

### 2022-12-15 2021-Refinement/paper/refinementarticleV2.tex
new refinement
work on refinement
V2

\usepackage{tikz,mathpartir}
\usepackage{mathtools}
\usepackage{amsmath,amssymb,mathabx}
%\usepackage{unicode-math}
\usepackage{tikz}
\usetikzlibrary{external}
\usepackage{hyperref}
\usepackage{nccrules}
\usepackage{comment}

@@

\DeclareMathOperator{\dom}{dom}
\newcommand{\rightarrowdbl}{\rightarrow\mathrel{\mkern-14mu}\rightarrow}

@@

\newcommand{\symb}[1]{\makebox{\it #1}}
\input{common.tex}
\input{preamble.tex}
\input{tikzmacros.tex}

@@

\begin{definition}[Hole-equal simulation]
Consider two OA with identical set of holes:  \(\OAg[1]\) and \(\OAg[2]\) where  \(J_1 = J_2\), the relation on configurations \(R: S_1 \times S_2 \to \rformulas[V_1 \uplus V_2]\) is a hole-\(\symb{equal}\) simulation from $S_1$ to $S_2$ if the following conditions hold: 
\item[(1)] \(\sigma_{01} \uplus \sigma_{02} \vdash R\mpar{s_{01}, s_{02}}\)
\item[(2)] \(\forall \mpar{s_1, s_2} \in S_1 \times S_2,\)\vspace{-8pt}
\noindent\begin{multline*}
\forall \mpar{s_1, s_2} \in S_1 \times S_2, \bigsymb{\forall} \mpar{t_1, t_2} = \mpar{\OTx{1}{}{1}{1}, \OTx{2}{}{2}{2}} \in \fOT{s_1} \times \fOT{s_2} \\ 
\everymath{\displaystyle}\begin{array}{l}
V_1 \uplus V_2 \uplus \fvars{t_1} \uplus \fvars{t_2}  \vdash R\mpar{s_1, s_2} \wedge g_2 \implies\\
V_1 \uplus V_2 \uplus \fvars{t_{1x}} \uplus \fvars{t_2}\vdash  \bigsymb{\bigexists} 
\mpar{t_{1x} \!=\! \OTx{1}{x}{1x}{1x}}^{x \in X} \operatorname*{\bigsymb{\bigvee}}_{x \in X}
\mpar{\begin{array}{l}
			\alpha_{1x} = \alpha_{2} \wedge \bigwedge_\subbox{j \in J'_{2}} \beta_{1xj} = \beta_{2j} \nwedge\\[12pt]
			 g_{1x} \wedge R\mpar{s'_{1x}, s'_{2}}\psubst{\psi_{1x} \uplus \psi_{2}}
		\end{array}} 
\end{array} 
\end{multline*}
\item [(3)] The deadlock reduction:
\begin{multline*}
\forall \mpar{s_1, s_2} \in S_1 \times S_2, V_1 \uplus V_2  \fvars{t_2} \vdash R\mpar{s_1, s_2} \wedge  \bigvee_\subbox{t_1 \in \fOT{s_1}} \fguard{t_1} \implies \bigvee_\subbox{t_2 \in \fOT{s_2}} \fguard{t_2}  
\end{multline*}

@@

%\section{a recuperer si on se rend compte que c'est utile}
%When we will introduce refinements in Section \ref{sec:prelref}, setting an undefined variable will be considered a valid refinement, for instance a \(5\) bits register is a particular \(n\) bits register.

@@

>>>>

\usepackage{tikz,mathpartir}
\usepackage{mathtools}
\usepackage{amsmath,amssymb,mathabx}
%\usepackage{unicode-math}
\usepackage{tikz}
\usetikzlibrary{external}
\usepackage{hyperref}
\usepackage{nccrules}
\usepackage{comment}
\usepackage{macrospNets}

@@

%\DeclareMathOperator{\dom}{dom}
%\newcommand{\rightarrowdbl}{\rightarrow\mathrel{\mkern-14mu}\rightarrow}

@@

%\newcommand{\symb}[1]{\makebox{\it #1}}
\input{common.tex}
\input{preamble.tex}
\input{tikzmacros.tex}
%\input{macrospNets.sty}

@@

\begin{definition}[Hole-equal simulation]
Consider two OA with identical set of holes:  \(\OAg[1]\) and \(\OAg[2]\) where  \(J_1 = J_2\), the relation on configurations \(R: S_1 \times S_2 \to \rformulas[V_1 \uplus V_2]\) is a hole-\(\symb{equal}\) simulation from $S_1$ to $S_2$ if the following conditions hold: 
\item[(1)] \(\sigma_{01} \uplus \sigma_{02} \vdash R\mpar{s_{01}, s_{02}}\)
\item[(2)] \(\forall \mpar{s_1, s_2} \in S_1 \times S_2,\)\vspace{-8pt}
\noindent\begin{multline*}
 \bigsymb{\forall} {t_1} = \openrule
         {
           \beta_{1 j}^{j\in J}, g_1,\psi_1}
         {s_1 \OTarrow {\alpha_1} s_1'} \in \fOT{s_1} .\,~~
\bigsymb{\bigexists} 
\mpar{t_{2x} \!=\! \openrule
         {
           \beta_{2 x j}^{j\in J}, g_{2 x},\psi_{2 x}}
         {s_2 \OTarrow {\alpha_{2 x}} s_{2 x}'}
}^{x \in X}\in \fOT{s_2}.\,~~\\ 
\everymath{\displaystyle}\begin{array}{l}
V_1 \uplus V_2 \uplus  \fvars{t_1} \uplus \bigcup_x\fvars{t_{2x}}  \vdash
\\
 R\mpar{s_1, s_2} \wedge g_1 \implies
 \operatorname*{\bigsymb{\bigvee}}_{x \in X}
\mpar{\begin{array}{l}
			\alpha_{2x} = \alpha_{1} \wedge \bigwedge_\subbox{j \in J} \beta_{2xj} = \beta_{1j} \nwedge\\[12pt]
			 g_{2x} \wedge R\mpar{ s'_{1},s'_{2x}}\psubst{\psi_{2x} \uplus \psi_{1}}
		\end{array}} 
\end{array} 
\end{multline*}
\item [(3)] The deadlock reduction:
\begin{multline*}
\forall \mpar{s_1, s_2} \in S_1 \times S_2, V_1 \uplus V_2  \vdash R\mpar{s_1, s_2} \wedge \lnot( \bigvee_\subbox{t_1 \in \fOT{s_1}} \fguard{t_1} )\implies \lnot( \bigvee_\subbox{t_2 \in \fOT{s_2}} \fguard{t_2}  )
\end{multline*}

@@

@@

\section{Proof of Transitivity for Refinement}\label{sec:proof-transitivity}
\begin{quote}
If $\wrel{A_1}{A_2}{H}$ and $\wrel{A_2}{A_3}{H'}$, then $\wrel{A_1}{A_3}{H\cap H'}$.
\end{quote}
\proof 
If $\wrel{A_1}{A_2}{H}$ then there is $R_{12}$ a relation between states
of $A_1$ and of $A_2$;  If $\wrel{A_2}{A_3}{H'}$ then there is $R_{23}$ a relation between states of $A_2$ and of $A_3$. We build a relation between
 states of $A_1$ and of $A_3$ as follows:  for each pair of states $s_1$, $s_3$, for each state $s_2$ such that $R_{12}$ relates $s_1$ and $s_2$, and $R_{23}$ relates $s_2$ and $s_3$.
Let $R_{13}$ be the relation:\\
  \[R_{13}(s_1,s_3)=\bigvee_{s_2\in S_2}\mpar{R_{12}(s_1,s_2) \land R_{23}(s_2,s_3) } \]

We will show that $\wrel{A_1}{A_3}{H\cap H'}$ by exhibiting  $R_{13}$ as a hole-tracking simulation of $A_1$ by  $A_3$.

We have to prove that the relation $R_{13}$ satisfies the three conditions of the definition of a refinement of open automata.
\begin{enumerate}
\item Firstly, we have to $R_{13}$ satisfies initial configurations:
\[\sigma_{01} \uplus \sigma_{03} \vdash R_{13} \mpar{s_{01}, s_{03}}\]
By knowing that substitutions only have an effect on the variables of the open automaton they belong to, they also produce terms containing only variables of the open automaton they belong to. We have:
\begin{align*}
\mpar{\sigma_{01} \uplus \sigma_{02} \vdash R_{12}\mpar{s_{01}, s_{02}}} \wedge \mpar
{\sigma_{02} \uplus \sigma_{03} \vdash R_{23}\mpar{s_{02}, s_{03}}}&\implies\\
R_{12}\mpar{s_{01},s_{02}}\psubst{\sigma_{01} \uplus \sigma_{02}} \wedge R_{23}\mpar{s_{02}, s_{03}}\psubst{\sigma_{02} \uplus \sigma_{03}}&\implies\\
R_{12}\mpar{s_{01},s_{02}}\psubst{\sigma_{01} \uplus \sigma_{02} \uplus \sigma_{03}} \wedge R_{23}\mpar{s_{02}, s_{03}}\psubst{\sigma_{01} \uplus \sigma_{02} \uplus \sigma_{03}}&\implies\\ 
\underbrace{R_{12}\mpar{s_{01},s_{02}}\wedge R_{23}\mpar{s_{02}, s_{03}}}_{\implies R_{13} \mpar{s_{01}, s_{03}}}\psubst{\sigma_{01} \uplus \sigma_{02} \uplus \sigma_{03}} %&\implies\\
 \end{align*}
Since $\sigma_{02}$ has no effect on variables of $s_{01}$ and $s_{03}$ thus we get the expected result.

\item \TODO{Ref to JLAMP 2023?}

\item Lastly, we have to prove the satisfaction of the deadlock reduction condition:
According to the relation between A2 and A3 we have for all

%\begin{multline*}
%\forall \mpar{s_1, s_2} \in S_1 \times S_2, V_1 \uplus V_2  \vdash R\mpar{s_1, s_2} \wedge \lnot( %\bigvee_\subbox{t_1 \in \fOT{s_1}} \fguard{t_1} )\implies \lnot( \bigvee_\subbox{t_2 \in \fOT{s_2}} %\fguard{t_2}  )
%\end{multline*}

\end{enumerate}
%\section{a recuperer si on se rend compte que c'est utile}
%When we will introduce refinements in Section \ref{sec:prelref}, setting an undefined variable will be considered a valid refinement, for instance a \(5\) bits register is a particular \(n\) bits register.
